The state of knowledge for marine invertebrate biodiversity in the continental US

### 2018-01-19 manuscript.Rmd
[fix #87] write about influence of missing geo coordinates

To assess the level of digitization of the fauna, we compared the species included in iDigBio and OBIS with the comprehensive checklists for molluscs [@turgeon1998] and crustaceans [@mclaughlin2005] that are available for the United States. For molluscs, of the `r format(itis_stats$moll_nspp, big.mark=",")` species included in the checklist, `r itis_stats$prop_moll_idig`% are in iDigBio and `r itis_stats$prop_moll_obis`% are in OBIS. For crustaceans, of the `r format(itis_stats$crust_nspp, big.mark=",")` species included in the checklist, `r itis_stats$prop_crust_idig`% are in iDigBio and `r itis_stats$prop_crust_obis`% are in OBIS. Given that molluscs and crustaceans are among the most sampled taxa, and that they have received more taxonomic attention than other phyla (see below), these percentages offer a conservative estimate for the fraction of the fauna that is available digitally. Therefore, using a conservative estimate of a 60% digitization level of the fauna, we can estimate that there is at least 29,000 species of marine invertebrates in US waters. The higher overlap between the comprehensive checklist for  crustaceans and OBIS is attributable to occurences of planktonic species that are more represented in OBIS data than in museum collections.

@@

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for those that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_eez*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_eez, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, is needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems. Similarly, only a small fraction of the marine invertebrates records from instiutions on the Pacific has been digitized.

>>>>

To assess the level of digitization of the fauna, we compared the species included in iDigBio and OBIS with the comprehensive checklists for molluscs [@turgeon1998] and crustaceans [@mclaughlin2005] that are available for the United States. For molluscs, of the `r format(itis_stats$moll_nspp, big.mark=",")` species included in the checklist, `r itis_stats$prop_moll_idig`% are in iDigBio and `r itis_stats$prop_moll_obis`% are in OBIS. For crustaceans, of the `r format(itis_stats$crust_nspp, big.mark=",")` species included in the checklist, `r itis_stats$prop_crust_idig`% are in iDigBio and `r itis_stats$prop_crust_obis`% are in OBIS. Given that molluscs and crustaceans are among the most sampled taxa, and that they have received more taxonomic attention than other phyla (Fig.  \@ref(fig:idigbio-through-time)), these percentages offer a conservative estimate for the fraction of the fauna that is available digitally. Therefore, using a conservative estimate of a 60% digitization level of the fauna, we can estimate that there is at least 29,000 species of marine invertebrates in US waters. The higher overlap between the comprehensive checklist for  crustaceans and OBIS is attributable to occurences of planktonic species that are more represented in OBIS data than in museum collections.

The faunistic list we assembled represent a subset of the species that have been digitized, because only records that are associated with geographic coordinates are included. All records published by OBIS have geographic coordinates associated with them. To estimate how records with missing geographic coordinates in iDigBio affect the completeness of our faunistic list, we queried iDigBio using the comprehensive lists for molluscs and crustaceans, and calculated the proportion of species with no iDigBio records within the US EEZ, and which had "United States" or "USA" in the country field, but had no geographic coordinates associated with them. These species may occur within the limits of the US EEZ but would not be returned in the queries used to assemble the faunistic list. We found that `r format_output(itis_stats$missing_geo_moll$prop_us_no_geo * 100)`% ($n=$ `r format(itis_stats$missing_geo_moll$n_spp, big.mark=",")`) of the species included in the comprehensive checklist for molluscs have records in iDigBio that place them in the United States based on the "Country" field, but do not have geographic coordinates associated with them that would ensures they fall within the limits of the EEZ. This percentage decreases to `r format_output(itis_stats$missing_geo_moll$prop_us_no_geo_no_obis * 100)`% once we exclude species that have informed geographic coordinates in OBIS. For crustaceans, these proportions are lower at `r format_output(itis_stats$missing_geo_crust$prop_us_no_geo * 100)`% and `r format_output(itis_stats$missing_geo_crust$prop_us_no_geo_no_obis * 100)`% ($n=$ `r format(itis_stats$missing_geo_crust$n_spp)`) respectively. This means ...

@@

Missing geographic coordinates reduces the overlap between the list of species from the regional list and the databases. Using iDigBio, we found that for the `r format(not_really_in_database$idig_gom_n_spp_not_in_db, big.mark=",")` species that do not appear within the geographic range of the Gulf of Mexico but are on the list, `r format_output(not_really_in_database$idig_gom_p_spp_not_in_db*100)`% are completely absent from iDigBio, while for the `r format(not_really_in_database$idig_gom_n_spp_no_coords, big.mark=",")` species that do appear in iDigBio, `r format_output(not_really_in_database$idig_gom_p_spp_no_coords*100)`%  do not have geographic coordinates associated with them, and for those that do have geographic coordinates, `r format_output(not_really_in_database$idig_gom_p_spp_within_eez*100)`% ($n=$  `r format(not_really_in_database$idig_gom_n_spp_within_eez, big.mark=",")`) fall within US waters. Additional digitization of collections, including the capture of their geographic coordinates information, is needed to improve biodiversity knowledge at regional scales. For instance, one can expect that the digitization coverage for the Gulf of Mexico will greatly increase once records from the Florida Wildlife Research Institute collections (St Petersburg, FL) become integrated into biodiversity information systems. Similarly, only a small fraction of the marine invertebrates records from institutions located on the Pacific coast has been digitized, as indicated by the paucity of Pacific coast collections listed in Table \@ref(tab:records-table).
